---
title: "P8105 Homework I"
date: "2022-09-27"
output:
  pdf_document: default
  html_document: default
---

```{r, message=FALSE}
# read and name data
data("penguins", package = "palmerpenguins")

# load the packages necessary for submission to knit.
library(tidyverse)

```

## Problem 1

The database _penguins_ has `r ncol(penguins)` variables and  `r nrow(penguins)` observations. Here's a code chunk that creates an overall summary table, including the names of all the variables. As we can see in the table, the mean **flipper length** is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

```{r, message=FALSE}

## create overall summary table
library(gtsummary)
penguins %>%
  tbl_summary(missing_text = "(Missing)", # counts missing values
              statistic = list(all_continuous() ~ "{mean} ({sd})"), #continuous variables
              label = list(sex ~ "Sex", # label variables
                           species ~"Species",
                           island ~ "Island",
                           year ~ "Year",
                           body_mass_g ~ "Body Mass",
                           flipper_length_mm ~ "Flipper Length",
                           bill_depth_mm ~ "Bill Depth",
                           bill_length_mm ~ "Bill Length" )) %>%
  bold_labels() %>%
  italicize_levels()
```

We now want to create a summary table by species. As we can see in the table, birds from the Gentoo species have, on average, lager flippers: 

```{r, message=FALSE}

# create summary table by species 
penguins %>%  
  tbl_summary(by = "species",   # stratify by species
              missing_text = "(Missing)",
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    label = list(sex ~ "Sex", # label variables
                 species ~"Species",
                 island ~ "Island",
                 body_mass_g ~ "Body Mass",
                 flipper_length_mm ~ "Flipper Length",
                 bill_depth_mm ~ "Bill Depth",
                 bill_length_mm ~ "Bill Length" )) %>%
  bold_labels() %>% 
  italicize_levels() 

```


We can also describe this data with plots.

```{r, message=FALSE}
library(ggplot2)

#create scatterplot

f1 <- ggplot(penguins) +    
      aes( x= flipper_length_mm, y= bill_length_mm,color=species) +
      geom_point () +
      labs(title="Flipper Length vs Bill Depth", x = "Flipper Length", y = "Bill Length") +
      theme_minimal()
                    
f1 
    png("f1.png")      

```


## Problem 2

Step 1 - Create a data frame:

```{r, message=FALSE}

# Create data frame 

df <- tibble(
        random_sample = rnorm(10),
        logical_vector = random_sample > 0,
        character_vector = c("a", "b","c","d","e", "f", "g", "h","i", "j"),
        factor_vector = factor(c("low","medium","high","low","medium","high","low","medium","high","low"))
        )
```

Step 2 - Take the mean of each variable in your dataframe:

```{r, message= FALSE, warrning = FALSE, eval = FALSE}

mean(df %>% pull(random_sample)) # Works 
mean(df %>% pull(logical_vector)) # Works
mean(df %>% pull(character_vector)) # Does not work, argument is not numeric or logical
mean(df %>% pull(factor_vector)) # Does not work, argument is not numeric or logical

```

Step 3 - Convert variables from one type to another and calculate the mean:

```{r,message= FALSE, warrning = FALSE, eval= FALSE}

# Code chunk that applies the as.numeric function to the logical, character, and factor variables 

new_1 <-as.numeric(df %>% pull(logical_vectir))
mean(numeric_log_var) #True is converted to 1 and False is converted to 0

new_2 <- as.numeric(df %>% pull(character_vector))
mean(new_2) #The vector is now numeric. We can now calculate the mean. 

new_3 <- as.numeric(df %>% pull(factor_vector))
mean(new_3) ##The vector is now numeric. We can now calculate the mean.

```