---
title: "P8105 Homework I"
date: "2022-09-27"
output:
  html_document: default
  pdf_document: default
---

```{r, message=FALSE}
# read and name data
data("penguins", package = "palmerpenguins")

# load the packages necessary for submission to knit.
library(tidyverse)

```

## Problem 1

The database _penguins_ has `r ncol(penguins)` variables and  `r nrow(penguins)` observations. Here's a code chunk that creates an overall summary table, including the names of all the variables. As we can see in the table, the mean **flipper length** is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

```{r, message=FALSE}

## create overall summary table
library(gtsummary)
penguins %>%
  tbl_summary(missing_text = "(Missing)", # counts missing values
              statistic = list(all_continuous() ~ "{mean} ({sd})"), #continuous variables
              label = list(sex ~ "Sex", # label variables
                           species ~"Species",
                           island ~ "Island",
                           year ~ "Year",
                           body_mass_g ~ "Body Mass",
                           flipper_length_mm ~ "Flipper Length",
                           bill_depth_mm ~ "Bill Depth",
                           bill_length_mm ~ "Bill Length" )) %>%
  bold_labels() %>%
  italicize_levels()
```

We now want to create a summary table by species. As we can see in the table, birds from the Gentoo species have, on average, lager flippers: 

```{r, message=FALSE}

# create summary table by species 
penguins %>%  
  tbl_summary(by = "species",   # stratify by species
              missing_text = "(Missing)",
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    label = list(sex ~ "Sex", # label variables
                 species ~"Species",
                 island ~ "Island",
                 body_mass_g ~ "Body Mass",
                 flipper_length_mm ~ "Flipper Length",
                 bill_depth_mm ~ "Bill Depth",
                 bill_length_mm ~ "Bill Length" )) %>%
  bold_labels() %>% 
  italicize_levels() 

```



We can also describe this data with plots.


## Problem 2
